Selection of Variables and Functional Forms in Multivariable Analysis

### 2025-04-16 index.qmd
Update index.qmd
Update index.qmd content
Update index.qmd with TG2 content

---
title: "Initial Data Analysis"
---

@@

To improve awareness of Initial Data Analysis (IDA) as an important part of the research process and to provide guidance on conducting IDA in a systematic and reproducible manner.

## Key principle

IDA does not look at associations of covariates with the outcome. The aim is to provide reliable knowledge about the data to ensure transparency and integrity of preconditions to conduct appropriate statistical analyses and correct interpretation of the results to answer pre-defined research questions.

## Definition and IDA Framework

Initial Data Analysis (IDA) consists of all steps performed on the data of a study typically between the end of the data collection/entry and start of those statistical analyses that address research questions. 
A systematic plan of IDA activities (IDAP) should be incoporated in the statistical analysis plan (SAP). Ideally, IDA should already be performed during ongoing data collections to detect and deal with data issues as early as possible.

![](assets/idaframework.png)

@@

Topic group 3 is a member of the [STRATOS](https://stratos-initiative.org) Initiative (STRengthening Analytical Thinking for Observational Studies) which is a large collaboration of experts in many different areas of biostatistical research. Ongoing research, discussions and activities within STRATOS are conducted in nine [topic groups](https://stratos-initiative.org/en/groups) and several cross-cutting [panels](https://stratos-initiative.org/en/panels).

![Conference BIRS 2016](assets/Stratos_group2016.jpg){fig-align="center"}

>>>>

---
title: "Selection of Variables and Functional Forms in Multivariable Analysis"
---

@@

Derive guidance for variable and function selection  in multivariable analysis.

## Focus

The main focus of TG2 is to identify influential variables and gain insight into their individual and joint relationship with the outcome. Two of the (interrelated) main challenges are selection of variables for inclusion in a multivariable explanatory model, and choice of the functional forms for continuous variables

@@

Topic group 2 is a member of the [STRATOS](https://stratos-initiative.org) Initiative (STRengthening Analytical Thinking for Observational Studies) which is a large collaboration of experts in many different areas of biostatistical research. Ongoing research, discussions and activities within STRATOS are conducted in nine [topic groups](https://stratos-initiative.org/en/groups) and several cross-cutting [panels](https://stratos-initiative.org/en/panels).

![2nd General Meeting of the STRATOS members at the Banff International Research Station (BIRS), Banff, Canada.](assets/STRATOS_Banff_second_meeting.png){fig-align="center"}
